---
title: ''
author: ''
date: ''
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### HW 4

``` {r readData , error = T,echo = FALSE}
#Load data & packages
library(faraway)
data(teengamb)
#head(teengamb)

#sex (levels(teengamb$test) <- c("Male","Female"))
# Change 0 to male, and 1 to female
teengamb$sex[teengamb$sex == 1] <- "FEMALE"
teengamb$sex[teengamb$sex == 0] <- "MALE"

# Change sex from quantitative to categorical through factor
teengamb$sex <- factor(teengamb$sex)
```

After running a linear model with all predictors present (sex,status, income, verbal), we get insignificant values for coefficients status & verbal. As a result, an F test is run to understand whether dropping these predictors from the analysis would be beneficial. Indeed, the F test impedes us to reject "H0: Bstatus = Bverbal=0" confirming their lack of need. Hence, our final model regresses gamble only on **sex** and **income**. 

```{r featureSelection, echo=FALSE}
lmod= lm(gamble ~ sex+status+income+verbal, data= teengamb)
lmod_siv = lm(gamble ~ sex+income, data= teengamb)
#F test to check whether Bstatus==0
anova(lmod_siv, lmod)
#print('Model Selected')
summary(lmod_siv)
```


1. One can check for constant variance by plotting residuals agains fitted values. For a better resolution one can take the square root of the abs(residuals).  Clearly, constant variance assumption is not valid. 
```{r constantVariance, echo=FALSE,  out.height='75%', out.width='61%', fig.align='center'}

#Check for constant variance by plotting residuals (y) vs fitted (x)
par(mfrow=c(1,2))

#Plot residuls against fitted  
plot(fitted(lmod_siv),residuals(lmod_siv), xlab= 'fitted values', ylab='residuals', main= expression(Residuals-"vs"-Fitted))
abline(h=0)

#Plot sqrt(abs(residuls)) double resolution  
plot(fitted(lmod_siv), sqrt(abs(residuals(lmod_siv))),xlab= 'fitted values',ylab=expression(sqrt(residuals)) , main= expression(sqrt(Residuals)-"vs"-Fitted))
```
> Furthermore, the non constant variance can be checked by looking at the strenght of the relationship between residuals and fitted. In a constant variance scenario, the R^2 should be zero.

```{r rsquared}
lmod_res_fit= lm(sqrt(abs(residuals(lmod_siv))) ~ fitted(lmod_siv))
summary(lmod_res_fit)$r.squared
```
> In order to "fix" this broken assumption, a transformation is applied to the response (sqrt). The constant variance does indeed check out now, as verified  by the residuals vs fitted values plot. 

```{r transformeResponse, echo = FALSE, out.height='50%', out.width='50%', fig.align='center'}
#Transform
t_lmod_siv= lm(sqrt(gamble) ~ sex+income, data= teengamb)

#Check for constant variance one more
plot(fitted(t_lmod_siv),residuals(t_lmod_siv), xlab='Fitted', ylab = 'Residuals',main = expression(lm(sqrt(gamble) ~ sex+income)))
abline(h=0)
print('The model with a transformed response')
summary(t_lmod_siv)
```

2. Normality can be checked through QQ plot (residuals/standardized), histogram , or Shapiro-Wilk (H0: residuals are normal) test. All graphical and numerical methods suggest normaility is satisfied.

```{r normality, echo=FALSE ,out.height='75%', out.width='75%', fig.align='center' }

par(mfrow=c(1,3))

#Normality is checked throguh QQ-plot 
qqnorm(residuals(t_lmod_siv),ylab="Residuals",main="QQ plot Residuals")
qqline(residuals(t_lmod_siv))

#Standardized residuals
qqnorm(rstandard(t_lmod_siv),ylab="Standardized Residuals",main='QQ plot Std. Res')
abline(0,1)

#Histogram
hist(residuals(t_lmod_siv),xlab="Residuals",main="Histogram of Residuals")

#Shapiro Wilk Test, Ho: residuals are normal (fail to reject)
shapiro.test(residuals(t_lmod_siv))
```

3. Leverage points are extreme values in the X space. One can check for such points through a half plot. Any point greater than 2p/n should be investigated further. In our case observations 33 & 42 are extremes.

```{r leveragePoints, echo=FALSE,out.height='75%', out.width='75%', fig.align='center'}
#Leverages are extreme values in the X spacv
hatv <- hatvalues(t_lmod_siv)
#Should sum to p (number of parameters)
#Wsum(hatv)
#head(hatv)
#Visualize through halfnormal plot , anything greater than 2p/n should be  considered more closely
hatv[which(hatv>2*3/length(hatv))]
halfnorm(hatv,ylab="Leverages", main='Half-normal plot for Leverages')

```

4. One can check for outliers by computing the studentized and apply Bonferroni's correction to find those points that do not fit the data. At an alpha level of 5% there are no apparent outliers given that maximum studentized residuals is smaller than the absolute value of the Bonferroni's critical value.

```{r outliers, echo=FALSE}
#Bonferroni's Correction , level = alpha/n (it is conservative) (ALPHA=0.05)
#Studentized
stud <- rstudent(t_lmod_siv)
print('Maximum Studentized Residual')
stud[which.max(abs(stud))]

#Conclusion: no outliers
#Reference: It is not worth the trouble of computing the outlier test p-value unless the studentized residual exceeds about 3.5 in absolute value.
```
```{r Bonferroni,}
#Bonferroni Critical Value (alph=0.05)
qt(.05/(47*2), t_lmod_siv$df.residual-1)
```
5. An influential point is a one whose removal from the data would cause a drastic change in the fit. One popular metric to identify such points is Cook's Distance.

```{r inlfuential,echo = FALSE, fig.align='center',out.height='75%', out.width='75%' }
#Cook statistic popular influential diagnostics

#half normal plots can be used to identify influential points
cook <- cooks.distance(t_lmod_siv)
halfnorm(cook,3,ylab="Cook’s distances", main='Influential Points')

#Should these values be dropped from the model? Does the model change significantly?
hat_lmod = lm(sqrt(gamble) ~ sex+income, data= teengamb, subset=(cook < max(
cook)))

#Plot, standardized residuals vs levarage
#plot(rstandard(t_lmod_siv), hatv)
```
> Droppiong the "influential" points from the model does not change the coefficinets, RSE, or R^2.

```{r dropMaxCook, echo=FALSE}
#Model
hat_lmod$call
#Coefficients
(summary(hat_lmod))$coefficients
#Residual Standard Error
(summary(hat_lmod))$sigma
#R^2
(summary(hat_lmod))$r.squared 
```

6. There is no ambiguous (i.e. non-linear) relationship between the predictors and the response.

```{r resPredRel, echo=FALSE, fig.align='center',out.height='75%', out.width='75%'}

par(mfrow=c(1,3))
plot(teengamb$income, residuals(t_lmod_siv), xlab = 'Income', ylab='Residuals', main='Residuals vs Predictor Structure',cex.main=0.85)
abline(h=0)

plot(teengamb$income, sqrt(teengamb$gamble), xlab = 'Income', ylab=expression(sqrt(gamble)), main='Response vs Predictor Structure', cex.main=0.85)

d = residuals(lm(sqrt(gamble) ~ sex,teengamb))
m = residuals(lm(income ~ sex,teengamb))
plot(m,d,xlab="Income residuals",ylab="Teengamb residuals",
     main='Partial Regression Structure',cex.main=0.85)
#coef(lm(d ~ m))
#coef(t_lmod_siv)
abline(0,coef(t_lmod_siv)['income'])
```
```{r partialRegression,fig.align='center',out.height='75%', out.width='75%', echo =FALSE, include=FALSE}
d = residuals(lm(sqrt(gamble) ~ sex,teengamb))
m = residuals(lm(income ~ sex,teengamb))
plot(m,d,xlab="Income residuals",ylab="Teengamb residuals",
     main='Partial Regression Structure',cex.main=0.8)
coef(lm(d ~ m))
coef(t_lmod_siv)
abline(0,coef(t_lmod_siv)['income'])
```
> This is the overall structure of the model (a gift from R). 

```{r structure, echo=FALSE,fig.align='center'}
par(mfrow=c(2,2))
plot(t_lmod_siv)
```



