---
title: ''
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### HW 6

1. Model  with *total* as the  response and *takers*, *ratio*, *salary*, and *expend* as predictors from dataset *sat*.

```{r data,echo=FALSE}
library(faraway)
data(sat)
```

a. Ordinary Least Squares, in this model **takers** seems to be the only significant (alpha=5%) predictor having a negative coefficient. All other predictors are insignificant with *ratio* having a negative coefficient and *salary* & *expend* having positive coefficients. The intercept is quite large and positive. The R^2 (0.82) is high & the RSE (32) is farily low. 

```{r ols,echo=FALSE}
ols_fit = lm(total ~ takers+ratio+salary+expend, data = sat)
summary(ols_fit)
```


b. Least Absolute Deviations, in this model **takers** & **ratio** seem to be both significant (alpha=5%) predictors having both negative coefficients. All other predictors are insignificant with *expend* having a negative coefficient and *salary* having a positive coefficient. The intercept is quite large and positive. The SE for *expend* is quite high resonating its insignificance.

```{r lib, include=FALSE, echo=FALSE}
library(quantreg)
```

```{r lad,echo=FALSE}
lad_fit = rq(total ~ takers+ratio+salary+expend, data = sat)
summary(lad_fit,se="nid")
```


c. Huber's Robust Regression, in this model **takers** seems to be the only significant (alpha=5%) predictor having a negative coefficient. All other predictors are insignificant with *ratio* having a negative coefficient and *salary* & *expend* having positive coefficients. The intercept is quite large and positive. The RSE (25) is farily low. 
```{r huber,echo=FALSE}
library(MASS)
library(sfsmisc)
huber_fit = rlm(total ~ takers+ratio+salary+expend, data = sat)
summary(huber_fit)
#P-values
f.robftest(huber_fit, var = "takers")
f.robftest(huber_fit, var = "ratio")
f.robftest(huber_fit, var = "salary")
f.robftest(huber_fit, var = "expend")
```

Model Comparison

**Ordinary least squares** and **Huber's Robust Regression** are closer to each other relative in their results. Both models have *takers* as the only significant predictor; both models share the same coefficients *signs* (takers (-), ratio(-),salary(+), expand(+)); both models have similar *intercept* (high & positive); both models share similar *RSE* (32 & 25); both models have similar *standard errors*. The models differ slightly in the *estimates of the coefficients*. The models p-values for **ratio** differ greatly- 26% vs 10%. 

**OLS/HRR** and **Least Absolute Deviations** are a bit further apart from each other relative in their results. Both models have *takers* as the a significant predictor-but LAD considers *ratio* as a significant predictor too; both models share the same coefficients *signs* (takers (-), ratio(-),salary(+))-but LAD considers *expend* negative in its sign (ols/HRR +); both models have similar *intercept* (high & positive); both models have similar *standard errors*. The models differ greatly in the *estimates of the coefficients*. The models p-values for **salary** differ greatly- on one side OLS/HRR  (49%,35%) and on the other LAD with 12%. 